Structure Learning for $\VAR(p)$ Models

### 2019-05-02 paper/paper.tex
Some figures and minor updates.

\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\VAR{\mathsf{VAR}}  % VAR(p) model
\def\B{\mathsf{B}}  % Filter B
\def\wtB{\widetilde{\B}}  % General filter B
\def\A{\mathsf{A}}  % Filter A
\def\H{\mathcal{H}}  % Hilbert space

@@

  \begin{enumerate}
    \item{$x_j \not\gc x_i$}
    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
    % \item{\hl{Wold $A(\tau)$ condition}}
    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
  \end{enumerate}
\end{theorem}

@@

  $(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
  (by definition) given by a form similar to equation
  \ref{eqn:ar_representation}:

@@

$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
$y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
$\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case
$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows
since $v_i(t) \perp \H_{t - 1}$ and
$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
  - 1}$

\hl{TODO: $(4) \implies (1)$}

@@

\begin{equation*}
  \begin{aligned}
    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
  \end{aligned}
\end{equation*}

@@

  Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e. $\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.  The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.  The second term is 0 since $j \not\in \anc{i}$.  The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.  For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, although similarly as before we can at least see immediately that

@@

  However, $k \ne \ell \implies \A_{k\ell}(z) = 0$ (i.e. $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.  This establishes the fact that $H_t^{(i)} \perp \H_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.  It is clear that $\H_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.

  \hl{TODO: Separate $H_t^i \perp \H_t^j \implies \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}

@@

  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.

  For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.  The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \not\pwgc i$, which will then establish the theorem.

@@

  which will imply (\hl{by lemma ???}) that $j \not\pwgc i$ and for which it is equivalent to show that

@@

It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
would hold, i.e. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
this is not the case in general, as different paths through $\gcg$ can
lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
fact, we do not even have $j \in \pa{i} \implies j \pwgc i$ (see
example \ref{ex:lag_cancellation}).

@@

we see that, since $v(t)$ is isotropic white noise, $1 \not\pwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
\end{example}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
    \label{fig:diamond_cancellation}
    \includegraphics[width=\linewidth]{example1.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
    \label{fig:lag_cancellation}
    \includegraphics[width=\linewidth]{example2.pdf}
  \end{subfigure}
\end{figure}

@@

\begin{align*}
  x_2(t) &= v_2(t) - ax_1(t - 1)\\
  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
  \implies x_3(t) &= v_2(t - 1) + v_3(t),
\end{align*}

and again $1 \not\pwgc 3$.
\end{example}

@@

\begin{corollary}
  \label{cor:parent_corollary}
  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
\end{corollary}
\begin{proof}
  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
  acyclic $i \not\in \anc{j}$, and by proposition
  \ref{prop:ancestor_properties} there is some
  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
  confounder, a contradiction.
\end{proof}
% This is a direct proof
  % \begin{proof}
%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
%   contradiction $i \in \anc{j}$.

@@

\begin{proposition}
  \label{prop:pwgc_anc}
  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \implies j \pwgc i$.
\end{proposition}
\begin{proof}
  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have

@@

\begin{corollary}
  \label{cor:gc_implies_pwgc}
  If $\gcg$ is a strongly causal DAG then $j \gc i \implies j \pwgc i$.
\end{corollary}

@@

    \BlankLine
    \caption{Pairwise Graph Recovery}
    \label{alg:pwgr}
    % \TitleOfAlgo{Pairwise Graph Recovery}
    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg = ([n], E)$.}
    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
      $k = 0$ \texttt{\# a counter used only for notation}}
    \BlankLine
    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \not\pwgc i\}$  \texttt{\# potential edges}\\
    \While{$S_k \ne \emptyset$}{
      $\widehat{P}_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
      % $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
      $S_{k + 1} \leftarrow S_k \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\

@@

      \For{$r = 0, \ldots, k$}
      {
        $D_{kr} \leftarrow \{(i, j) \in \widehat{P}_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
      }
      $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\

@@

We prove the theorem by establishing the correctness of algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.  The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^k\widehat{P}_{k - r}$ and can be determined by a pairwise test.

\begin{example}
  That we need to proceed backwards through $\widehat{P}_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k \widehat{P}_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.  Moreover, the same example shows that if simply $D_k = \{(i, j) \in \widehat{P}_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.

@@

  \begin{figure}
    \centering
    \label{fig:example_fig3}
    \caption{Example graph for Algorithm \ref{alg:pwgr}}
    \includegraphics[width=0.5\linewidth]{hand_drawn_example3.jpg}
  \end{figure}
\end{example}

@@

\begin{proof}
% 1. Dispense with trivialities
% 2. Show that hat{P}_k is nonempty
% 3. Show that C_k is nonempty
% 4. Show that edges in D_k are actual edges in G
% 5. show that each edge incident on C_k is in D_k
% 6. Induction

% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define

  \begin{equation*}
    A_k = \{i \in [n]\ |\ \dist{i}{\widehat{P}_0} = k\},
  \end{equation*}

  where

  \begin{equation*}
    \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
  \end{equation*}

  We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $\widehat{P}_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.  If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $\widehat{P}_0$.  That is, $\bigcup_{\ell \le n}A_\ell = [n]$.

  % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.

  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.  To quickly dispense with some trivialities: the set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.  Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.

  Continuing with the non-trivial case: if $\widehat{P}_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.  In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.  We must have $j \in C_0$, otherwise there would be a path of length $2$ from $\widehat{P}_0$ to $j$.  Moreover, since $\exists i \in \widehat{P}_0$ s.t. $i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.  Therefore $A_1 \subseteq D_{00}$.  \hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}

  Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).  Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).  Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.  This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.  So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.

  Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.

% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
%   On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.  Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).  Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).  Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.  So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.  Suppose now that for some $j \in C_0$ there is an $i$ s.t. $(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.  Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in \widehat{P}_0$ then $j \notin C_0$ and if $i \in \widehat{P}_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.

%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.

>>>>

\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\VAR{\mathsf{VAR}}  % VAR(p) model
\def\B{\mathsf{B}}  % Filter B
\def\wtB{\widetilde{\B}}  % General filter B
\def\A{\mathsf{A}}  % Filter A
\def\H{\mathcal{H}}  % Hilbert space

@@

  \begin{enumerate}
    \item{$x_j \ngc x_i$}
    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
    % \item{\hl{Wold $A(\tau)$ condition}}
    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
  \end{enumerate}
\end{theorem}

@@

  $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
  (by definition) given by a form similar to equation
  \ref{eqn:ar_representation}:

@@

$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
$y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
$\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case
$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows
since $v_i(t) \perp \H_{t - 1}$ and
$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
  - 1}$

\hl{TODO: $(4) \Rightarrow (1)$}

@@

\begin{equation*}
  \begin{aligned}
    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
    \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
  \end{aligned}
\end{equation*}

@@

  Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e. $\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.  The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.  The second term is 0 since $j \not\in \anc{i}$.  The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.  For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, although similarly as before we can at least see immediately that

@@

  However, $k \ne \ell \Rightarrow \A_{k\ell}(z) = 0$ (i.e. $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.  This establishes the fact that $H_t^{(i)} \perp \H_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.  It is clear that $\H_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.

  \hl{TODO: Separate $H_t^i \perp \H_t^j \Rightarrow \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}

@@

  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \npwgc i$.

  For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.  The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \npwgc i$, which will then establish the theorem.

@@

  which will imply (\hl{by lemma ???}) that $j \npwgc i$ and for which it is equivalent to show that

@@

It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
would hold, i.e. $j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately,
this is not the case in general, as different paths through $\gcg$ can
lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
fact, we do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see
example \ref{ex:lag_cancellation}).

@@

we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
\end{example}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\linewidth]{example1.pdf}
    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
    \label{fig:diamond_cancellation}
    \centering{$j \in \anc{i} \nRightarrow j \pwgc i$}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\linewidth]{example2.pdf}
    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
    \label{fig:lag_cancellation}
    \centering{$j \in \pa{i} \nRightarrow j \pwgc i$}
  \end{subfigure}
\end{figure}

@@

\begin{align*}
  x_2(t) &= v_2(t) - ax_1(t - 1)\\
  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
\end{align*}

and again $1 \npwgc 3$.
\end{example}

@@

\begin{corollary}
  \label{cor:parent_corollary}
  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
  \textit{alternatives}, that is $i \pwgc j \Rightarrow j \notin \anc{i}$.
\end{corollary}
\begin{proof}
  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
  acyclic $i \not\in \anc{j}$, and by proposition
  \ref{prop:ancestor_properties} there is some
  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
  confounder, a contradiction.
\end{proof}
% This is a direct proof
  % \begin{proof}
%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
%   contradiction $i \in \anc{j}$.

@@

\begin{proposition}
  \label{prop:pwgc_anc}
  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
\end{proposition}
\begin{proof}
  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have

@@

\begin{corollary}
  \label{cor:gc_implies_pwgc}
  If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
\end{corollary}

@@

    \BlankLine
    \caption{Pairwise Graph Recovery}
    \label{alg:pwgr}
    % \TitleOfAlgo{Pairwise Graph Recovery}
    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg$.}
    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
      $k = 0$ \texttt{\# a counter used only for notation}}
    \BlankLine
    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
    \While{$S_k \ne \emptyset$}{
      $P_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
      % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
      $S_{k + 1} \leftarrow S_k \setminus P_k$ \texttt{\# remove from } $S$\\

@@

      \For{$r = 0, \ldots, k$}
      {
        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
      }
      $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\

@@

We prove the theorem by establishing the correctness of algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.  The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^kP_{k - r}$ and can be determined by a pairwise test.

\begin{example}

  That we need to proceed backwards through $P_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k P_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.  Moreover, the same example shows that if simply $D_k = \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.

@@

  \begin{figure}
    \centering
    \label{fig:example_fig3}
    \caption{Example graph for Algorithm \ref{alg:pwgr}}
    \footnotesize{Black arrows indicate true parent-child
      relations.  Red dotted arrows indicate pairwise causality (due to
      ancestor relations), green dash-dotted arrow indicates
      bidirectional pairwise causality (due to the confounding node
      $1$).  Blue groupings indicate each $P_k$ in Algorithm
      \ref{alg:pwgr}.  Repetition is for aesthetics.}

    \begin{subfigure}[b]{0.45\textwidth}
      \includegraphics[width=\linewidth]{example_algorithm.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
      \includegraphics[width=\linewidth]{example_algorithm2.pdf}
    \end{subfigure}
  \end{figure}
\end{example}

@@

\begin{proof}
  Denote the edges of $\gcg$ by $\mathcal{E}$, we endevour to show
  that upon the termination of Algorithm \ref{alg:pwgr} we have
  $E = \mathcal{E}$.  On the first iteration of the algorithm
  ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.

  We will first dispense with some trivialities: if $n = 0$ or $n = 1$
  the algorithm is correct, returning on the first iteration with
  $E = \emptyset$.  Suppose then that $n > 1$.  The set
  $P_0$ is non-empty since $\gcg$ is a DAG and therefore
  there are nodes in $\gcg$ without parents, thence if
  $P_0 = \emptyset$ proposition
  \ref{prop:ancestor_properties} would be contradicted.  Now, if
  $P_0 = [n]$ then the graph necessarily contains no edges
  and we will correctly return $E = \emptyset$.

  Moving on from any trivial cases, notice that
  $\mathcal{E} \subseteq W_0$.  This follows since by Proposition
  \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
  Corollary \ref{cor:bidirectional_edge} there are no bidirectional
  edges in $\gcg$.

% 1. Dispense with trivialities
% 2. Show that hat{P}_k is nonempty
% 3. Show that C_k is nonempty
% 4. Show that edges in D_k are actual edges in G
% 5. show that each edge incident on C_k is in D_k
% 6. Induction

% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
  % Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define

  % \begin{equation*}
  %   A_k = \{i \in [n]\ |\ \dist{i}{P_0} = k\},
  % \end{equation*}

  % where

  % \begin{equation*}
  %   \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
  % \end{equation*}

  % We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $P_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.  If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $P_0$.  That is, $\bigcup_{\ell \le n}A_\ell = [n]$.

  % % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.

  % On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.  To quickly dispense with some trivialities: the set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.  Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.

  % Continuing with the non-trivial case: if $P_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $P_0$.  In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.  We must have $j \in C_0$, otherwise there would be a path of length $2$ from $P_0$ to $j$.  Moreover, since $\exists i \in P_0$ s.t. $i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.  Therefore $A_1 \subseteq D_{00}$.  \hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}

  % Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).  Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).  Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.  This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.  So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.

  % Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
